\frb: Properties of VLA Bursts and Implications for Overall FRB Population

### 2016-11-22 repeater-vladata.tex
draft circulated with acf figure

\begin{abstract}
The highly-dispersed, millisecond radio transients known as Fast Radio Bursts have recently emerged as a new class...

The discovery of repeating bursts from \frb\ has shown that at least some FRBs are not cataclysmic and opened potential for studying a homogenous sample of bursts...
Our recent coordinated campaign with the Very Large Array and Arecibo Observatory has made the first direct, arcsecond-scale localization of an FRB and unambiguously associated it with counterparts in other observations. This campaign detected nine bursts at the VLA from 2.5 to 3.5~GHz and many more at Arecibo, GBT, and Effelsberg. The coordination of these observations allow us to refine our picture of the physical processes in \frb.

With so many bursts, we can make the first reliable measures of burst flux distribution and temporal statistics...

The connection of \frb\ to the overall population...

\end{abstract}

@@

\frb\ has now been localized three orders of magnitude better than any other FRB and placed at a cosmological distance. Its lookback and luminosity distances are 746 and 972 Mpc \citep{planck15}, which are orders of magnitude larger than any other millisecond transient. This shows that FRBs:
\begin{itemize}
 \item are extremely luminous, 
 \item have a significant DM contribution from the IGM, and
 \item can be used to probe the IGM and their host galaxy.
\end{itemize}
The promise implied by the first reported FRB is now being realized \citep{2007Sci...318..777L}.

The cosmological distance for \frb\ could have wide-ranging implications for the FRB population as a whole. However, it has not been demonstrated that \frb\ is representative of the overall FRB population. In fact, the unambiguous repetition of its bursts is unique among all FRBs \citep{2015MNRAS.454..457P}, so it is natural to ask whether \frb\ is representative. An important first step is to demonstrate that the properties of \frb\ are consistent with the significant body of facts for the overall population \citep{2015MNRAS.451.3278M, 2016MPLA...3130013K}. The repeating nature of \frb\ provides us with several statistical tests we can use to test this connection.

We can also assume that \frb\ is representative and use it to constrain the physical processes at play in the overall FRB population. Although we now know that FRBs are luminous, it is not yet clear what process generates the radio bursts themselves \citep{2014PhRvD..89j3009K, 2014ApJ...785L..26L, 2016MNRAS.457..232C}. The simultaneous \frb\ observing campaign with the VLA, Arecibo, Effelsberg, GBT, and AMI gives a more complete picture of the spectral structure of FRB radio emission. FRB repetition also has strong implications for the number of FRB-generating systems in the universe \citep{2016MNRAS.458L..89C}.

@@

\section{Observations}
\begin{figure*}[t]
\begin{center}
\includegraphics[width=2\columnwidth]{timeline}
\caption{Summary of observing coverage during the multi-telescope observing campaign. Uncircled symbols show days with observations and circles show detections of \frb. *Only the coincident Arecibo burst is currently circled*. The black dashed line shows the VLA burst detections with simultaneous coverage with other telescopes.
\label{fig:multi}}
\end{center}
\end{figure*}

The data presented here were obtained from multiple programs and telescopes, but the central goal was to interferometrically localize \frb\ with the VLA. We ultimately settled on a strategy based on simultaneous observing between the VLA and Arecibo observatories with other observatories added on a best-effort basis. The hope was that Arecibo could be used to define times of bursts for which we could selectively search with the VLA. This would reduce the trials factor that limited sensitivity and potentially provide detections at multiple independent bands. Below, we summarize these observations, with a focus on those conducted simultaneous with VLA burst detections from \frb.

@@

Observations in August and Septmeber were searched in quasi real-time by a prototype version of \rf\footnote{See \url{http://realfast.io}.}. \rf\ is a real-time, fast imaging transient search system. The current, prototype runs on existing, CPU-based hardware of the VLA correlator backend, while the future \rf\ will run on a dedicated GPU cluster. The transient search pipeline software is called ``rtpipe''\footnote{See \url{https://github.com/caseyjlaw/rtpipe}} and is mostly written in Python. The transient search pipeline is described in \citet{2015ApJ...807...16L} and performs data calibration, flagging, dedispersion, and imaging. Images were formed for each integration at DMs of 0, 546, 556.9, 560, and 565 pc cm$^{-3}$. Each image with a pixel higher than 6.4$\sigma$ is saved to capture both real and thermal-noise candidates as a check of data quality. Gain calibration is read from the "telcal" system, which uses phase-only calibration on the previous gain calibrator.

Burst detections and localizations were made within hours of data being recorded. The transient search starts when data are recorded, but this prototype of \rf\ is a factor of several times slower than real-time, so we refer to the detection as "quasi real-time". When a candidate significance higher than 7.4$\sigma$, \rf\ generates a candidate visualization with an image and spectrum.

More detailed analysis, including improved calibration and localization, is conducted offline. Computational notebooks to reproduce the transient detection and localization can be found at \url{https://github.com/caseyjlaw/FRB121102}. Time cut-out visibility data are available at \url{https://doi.org/10.7910/DVN/TLDKXG}. Original visibility data are available under VLA program codes 16A-459 and 16A-496 and can be downloaded at \url{http://archive.nrao.edu}.

@@

4 observations during VLA bursts...

@@

\subsection{VLA Detections}

Figure \ref{fig:sgram} shows the spectrograms of all nine bursts detected by \rf.

@@

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.25\columnwidth]{sgram_57646.png}
  \includegraphics[width=0.25\columnwidth]{sgram_57648.png}
  \includegraphics[width=0.25\columnwidth]{sgram_57649.png}
 \end{minipage}
 \caption{Spectrograms (time vs frequency intensity maps) for all nine VLA bursts. Note that bursts are detected in 5~ms images generated from dedispersed visibilities. **NOT SURE WHY ONE FIG LARGER**
 \label{fig:sgram}}
\end{center}
\end{figure*}

\subsection{Multi-Telescope Burst Spectra}

See Figure \ref{fig:multi} for coverage of VLA bursts... Arecibo had coverage for 57648 VLA burst and sees a coincident burst. Effelsberg had simultaneous coverage for VLA bursts on 57648 and 57649, but no detections. AMI had simultaneous coverage for bursts on 57643, 57645, 57648, and 57649.

%\begin{table}
%\caption{Caption}
%\centering
%\begin{tabular}{llll}
%\hline
%Field       & RA          & Dec   & Lon. \\ \hline
%\end{tabular}
%\label{fields}
%\end{table}

\subsection{VLA Spectra}
After detection, we refined the analysis of each burst offline. We used a fine dispersion grid to find the best-fit DM for each burst.

Table?
57623 (DM$=561$ pc cm$^{-3}$)
57633, Scan 7 (DM$=554$ pc cm$^{-3}$)
57633, Scan 13 (DM$=559$ pc cm$^{-3}$)
57638 (DM$=554$ pc cm$^{-3}$)
57643 (DM$=559$ pc cm$^{-3}$)
57645 (DM$=572$ pc cm$^{-3}$)
57646 (DM$=573$ pc cm$^{-3}$)
57648 (DM$=559$ pc cm$^{-3}$)
57649 (DM$=552$ pc cm$^{-3}$).

Dispersion measure changes between bursts significant?

Burst spectra are shown in Figure \ref{fig:spec}...

@@

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.3\columnwidth]{spec_57646.png}
  \includegraphics[width=0.3\columnwidth]{spec_57648.png}
  \includegraphics[width=0.3\columnwidth]{spec_57649.png}
 \end{minipage}
\caption{Spectra of nine bursts seen by the VLA from 2.5 to 3.5~GHz.  **LABELS ALL SAME FREQ. SOME SHOULD START HIGHER FOR DROPPED CHANNELS**
 \label{fig:spec}}
\end{center}
\end{figure*}

@@

The VLA S-band recievers natively measure circular polarization, although observations did not include polarization calibration procedures. Crude constraints on circular polarization are possible by comparing the burst intensity in right and left-hand polarized data products. The apparent circular polarization fraction ($(RR-LL)/(RR+LL)$) for the most significant bursts are all less than 3\%. FRB121102 was located 2.3 arcmin away from pointing center, where systematic effects have been measured as large as 3\% (Perley et al 2016, VLA memo). We conclude that the S-band bursts seen by the VLA have a circular polarization of less than 3\%.

@@

The bursts seen by the VLA range in significance from 10 to 150$\sigma$. Offline analysis for flux scale...

Figure \ref{fig:logns} shows the flux distribution of the nine VLA bursts. Flatness...

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{logns}
\caption{Log N -- Log SNR distribution of real-time detections of FRB 121102.
\label{fig:logns}}
\end{center}
\end{figure}

Doubts were cast on the first FRB detection (``Lorimer burst'') due to its unusually high brightness. The lack of lower-significance detections suggested that this burst was unlikely to be part of any astrophysical population...

\subsection{Temporal Statistics}

Burst detections were made very inhomogenously though the roughly 63 hours of observing toward FRB 121102. Data quality are relatively high and RFI did not significantly impact sensitivity, so the inhomogeneous burst distribution not an observational artifact.

@@

\subsection{Connection to Overall FRB Population}

Flat luminosity distribution of FRB121102 and overall population

Does local (100 Mpc) population with flat luminosity distribution reproduce observed properties? Imagine also how log N/log S cut-offs and scattering biases the intrinsic into the observed distribution (Macquart and Johnston)

Discussion of ``red spectrum'' and \citet{2016MNRAS.458L..89C}. Bursts predict bursts therefore repetition constraints of other FRBs are likely weaker than claimed...

Intrinsic versus refractive scintillation

@@

@@

\subsection{Extending FRB 121102 to Other FRBs}

FRB 121102 has a DMhost of xx. Assuming other FRBs have a comparable DMhost, what can we infer about the population as a whole?

@@

Targeting known FRBs is optimal

Wide, low-sensitivity searches are best way to conduct a blind search

\cite{2016MNRAS.458L..89C}

>>>>

\begin{abstract}
The millisecond radio transients known as Fast Radio Bursts have recently emerged as a mysterious, new class of astrophysical transient. The discovery of repeating bursts from \frb\ has shown that at least some FRBs are not cataclysmic and opened potential for studying a FRB properties via a homogenous sample of bursts. With the recent localization of \frb with the Very Large Array has helped measure its distance and intrinsic properties of the bursting source and its multiwavelength associations. This localization was made with 9 bursts seen by the VLA in coordination the Arecibo, Effelsberg, and AMI-LA observatories. We present a detailed analysis of these bursts, including the first simultaneous detection of an FRB with multiple telescopes. We show that the burst spectra typically have a broad Gaussian shape on the scale of $\sim500$~MHz with fine spectral structure consistent with either scintillation or nanosecond temporal structure. We present the first burst flux distribution and temporal statistics for \frb and argue that it is consistent with the broader FRB population...
\end{abstract}

@@

\frb\ has now been localized three orders of magnitude better than any other FRB and placed at a cosmological distance. Its lookback and luminosity distances are 746 and 972 Mpc \citep{planck15}, which are orders of magnitude larger than any other millisecond transient. This shows that FRBs:
\begin{itemize}
 \item are extremely luminous, 
 \item have a significant DM contribution from the IGM, and
 \item can be used to probe the IGM and their host galaxy.
\end{itemize}
The promise implied by the first reported FRB \citep{2007Sci...318..777L} is now being realized.

The cosmological distance for \frb\ could have wide-ranging implications for the FRB population as a whole. However, it has not been demonstrated that \frb\ is representative of the overall FRB population. In fact, the repetition of its bursts is unique among all FRBs \citep{2015MNRAS.454..457P}, so it is natural to ask whether \frb\ is representative. An important first step is to demonstrate that the properties of \frb\ are consistent with the significant body of facts for the overall population \citep{2015MNRAS.451.3278M, 2016MPLA...3130013K}. The repeating nature of \frb\ provides us with several statistical tests we can use to test this connection.

We can also assume that \frb\ is representative and use it to constrain the physical processes at play in the overall FRB population. Although we now know that FRBs are luminous, it is not yet clear what process generates the radio bursts themselves \citep{2014PhRvD..89j3009K, 2014ApJ...785L..26L, 2016MNRAS.457..232C}. The simultaneous \frb\ observing campaign with the VLA, Arecibo, Effelsberg, GBT, and AMI-LA gives a more complete picture of the spectral structure of FRB radio emission. FRB repetition also has strong implications for the number of FRB-generating systems in the universe \citep{2016MNRAS.458L..89C}.

@@

\section{Observations}

\begin{figure*}[t]
\begin{center}
\includegraphics[width=2\columnwidth]{timeline}
\caption{Summary of observing coverage during the multi-telescope observing campaign. Uncircled symbols show days with observations and circles show detections of \frb. Only the coincident Arecibo burst is currently circled (**change?**). The black dashed line shows the VLA burst detections with simultaneous coverage with other telescopes.
\label{fig:multi}}
\end{center}
\end{figure*}

The data presented here were obtained from multiple programs and telescopes, but the central goal was to interferometrically localize \frb\ with the VLA. The observing strategy was to ensure simultaneous observing between the VLA and Arecibo observatories and add other observatories added on a best-effort basis. The hope was that Arecibo could be used to define times of bursts for which we could selectively search with the VLA. This would reduce the trials factor that limited sensitivity and potentially provide detections at multiple independent bands. Below, we summarize these observations, with a focus on those conducted simultaneous with VLA burst detections from \frb.

@@

Observations in August and Septmeber were searched by a prototype version of \rf\footnote{See \url{http://realfast.io}.}. \rf\ is a real-time, fast imaging transient search system. The current, prototype runs on existing, CPU-based hardware of the VLA correlator backend, while the future \rf\ will run on a dedicated GPU cluster. The transient search pipeline software is called ``rtpipe''\footnote{See \url{https://github.com/caseyjlaw/rtpipe}} and is mostly written in Python. The transient search pipeline is described in \citet{2015ApJ...807...16L} and performs data calibration, flagging, dedispersion, and imaging. Images were formed for each integration at DMs of 0, 546, 556.9, 560, and 565 pc cm$^{-3}$. Gain calibration is read from the "telcal" system, which uses phase-only calibration on the previous gain calibrator.

Burst detections and localizations were made within hours of data being recorded. The transient search starts when data are recorded, but this prototype of \rf\ is a factor of several times slower than real-time, so we refer to the detection as "quasi real-time". Each image with a pixel higher than 6.4$\sigma$ has some rudimentary data saved to capture both real and thermal-noise candidates as a check of data quality. For each image with a pixel higher than 7.4$\sigma$, \rf\ generates a candidate visualization with an image and spectrum. More detailed analysis, including improved calibration and localization, is conducted offline.

Computational notebooks to reproduce the transient detection and localization can be found at \url{https://github.com/caseyjlaw/FRB121102}. Time cut-out visibility data are available at \url{https://doi.org/10.7910/DVN/TLDKXG}. Original visibility data are available under VLA program codes 16A-459 and 16A-496 and can be downloaded at \url{http://archive.nrao.edu}.

@@

4 observations during VLA bursts... Upper limit on burst flux density limits any significant "afterglow". Deep image sensitivity is limited by confusion and is approximately equal to the flux density observed by VLA (CITE).

@@

\subsection{VLA Bursts}

\subsubsection{Dynamic Spectra}
Figure \ref{fig:sgram} shows the spectrograms of all nine bursts detected by \rf.

@@

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.25\columnwidth]{sgram_57646.png}
  \includegraphics[width=0.25\columnwidth]{sgram_57648.png}
  \includegraphics[width=0.25\columnwidth]{sgram_57649.png}
 \end{minipage}
 \caption{Spectrograms (time vs frequency intensity maps) for all nine VLA bursts. Note that bursts are detected in 5~ms images generated from dedispersed visibilities.
 \label{fig:sgram}}
\end{center}
\end{figure*}

We used a fine dispersion grid to find the best-fit DM for each burst. Table \ref{tab:spec} shows the best-fit DM values... Dispersion measure changes between bursts significant?

%\begin{table}
%\caption{Caption}
%\centering
%\begin{tabular}{llll}
%\hline
%Field       & RA          & Dec   & Lon. \\ \hline
%\end{tabular}
%\label{tab:spec}
%\end{table}

\subsubsection{Spectral modeling}
After detection, we refined the analysis of each burst offline. Burst spectra are shown in Figure \ref{fig:spec}...

@@

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.3\columnwidth]{spec_57646.png}
  \includegraphics[width=0.3\columnwidth]{spec_57648.png}
  \includegraphics[width=0.3\columnwidth]{spec_57649.png}
 \end{minipage}
\caption{Spectra of nine bursts seen by the VLA from 2.5 to 3.5~GHz.  **LABELS ALL SAME FREQ. SOME SHOULD START HIGHER FOR DROPPED CHANNELS**
%Table?
%57623 (DM$=561$ pc cm$^{-3}$)
%57633, Scan 7 (DM$=554$ pc cm$^{-3}$)
%57633, Scan 13 (DM$=559$ pc cm$^{-3}$)
%57638 (DM$=554$ pc cm$^{-3}$)
%57643 (DM$=559$ pc cm$^{-3}$)
%57645 (DM$=572$ pc cm$^{-3}$)
%57646 (DM$=573$ pc cm$^{-3}$)
%57648 (DM$=559$ pc cm$^{-3}$)
%57649 (DM$=552$ pc cm$^{-3}$).
\label{fig:spec}}
\end{center}
\end{figure*}

We fit a Gaussian model to each burst to estimate their characteristic width and peak flux... As shown in Table \ref{tab:spec}, the typical frequency scale is xx MHz...

\subsubsection{Spectral Autocorrelation}
The spectral autocorrelation is used to infer both intrinsic and scintillation properties \citep{}. Figure \ref{fig:acf} shows the spectral autocorrellation for the MJD 57633, scan 7 burst. Some marginal evidence of spectral correlation on the scale of the channel resolution (4~MHz)?

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{acf_57633_scan7}
\caption{
\label{fig:acf}}
\end{center}
\end{figure}

\subsubsection{Circular Polarization}
The VLA S-band recievers natively measure circular polarization, although observations did not include polarization calibration procedures. Crude constraints on circular polarization are possible by comparing the burst intensity in right and left-hand polarized data products. The apparent circular polarization fraction ($(RR-LL)/(RR+LL)$) for the most significant bursts are all less than 3\%. FRB121102 was located 2.3 arcmin away from pointing center, where systematic effects have been measured as large as 3\% (Perley et al 2016, VLA memo). We conclude that the S-band bursts seen by the VLA have a circular polarization of less than 3\%.

\subsection{Multi-Telescope Burst Spectra}
See Figure \ref{fig:multi} for coverage of VLA bursts... Arecibo had coverage for 57648 VLA burst and sees a coincident burst. No other coincident bursts seen. **Make figure of VLA/AO dynamic spectrum on 57648**

%\begin{figure}[htb]
%\begin{center}
%\includegraphics[width=0.9\columnwidth]{}
%\caption{
%\label{fig:}}
%\end{center}
%\end{figure}

@@

@@

The bursts seen by the VLA range in significance from 10 to 150$\sigma$. Table \ref{tab:spec} shows the estimated burst luminosity from spectral modeling...

Figure \ref{fig:logns} shows the flux distribution of the nine VLA bursts. Quantify flatness...

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{logns}
\caption{Log N -- Log SNR distribution of real-time detections of FRB 121102. **Update to luminosity distribution**
\label{fig:logns}}
\end{center}
\end{figure}

Compare luminosity distribution to flux distribution measured by \citep{2016ApJ...830...75V}...

\subsection{Temporal Statistics}
Burst detections were made very inhomogenously though the roughly 63 hours of observing toward FRB 121102. Data quality are relatively high and RFI did not significantly impact sensitivity, so the inhomogeneous burst distribution not an observational artifact.

@@

\subsection{Luminosity Distribution}
Doubts were cast on the first FRB detection (``Lorimer burst'') due to its unusually high brightness. The lack of lower-significance detections suggested that this burst was unlikely to be part of any astrophysical population. With more detections, it has become clear that the FRB population has a relatively flat flux distribution \citep{2016ApJ...830...75V, 2016arXiv161100458L}. This fact was demonstrated with yet another detection of an extremely bright FRB \citep{2016arXiv161105758R}.

Discuss comparison FRB 121102 luminosity distribution to that of FRB population...
Imagine physical log N/log S with cut-offs and scattering can bias the intrinsic into the observed distribution (Macquart and Johnston)...

\subsection{Repetition}
Discussion of ``red spectrum'' and \citet{2016MNRAS.458L..89C}. Bursts predict bursts therefore repetition constraints of other FRBs are likely weaker than claimed...

@@

Intrinsic versus refractive scintillation...

Fewer FRBs out there...

@@

@@

Repetition implies that targeting known FRBs is optimal.

Repetition and shallow luminosity distribution show that shallow and wide is the best way to blindly find FRBs.
